### 2018-07-02 R/manuscript_app.Rmd
finish code polishing

---
title: "manuscript - APP"
author: "Filip Dechterenko"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

## Load libraries

```{r load libraries, warning=F, message=F}

if (!require(tidyverse)) {install.packages("tidyverse"); library(tidyverse)}
if (!require(ez)) {install.packages("ez"); library(ez)}
if (!require(psyphy)) {install.packages("psyphy"); library(psyphy)}

if (!require(here)) {install.packages("here")}

theme_set(theme_bw(16))
source("tools.R")

data_pth <- here::here("..", "data")
plot_pth <- here::here("..", "plots")

if(!dir.exists(plot_pth)) {
  dir.create(plot_pth)
}
```

# Experiment 1a

```{r load data - exp 1a}
df_exp1a <- readRDS(file.path(data_pth,"exp1a","responses","responses_150330.rds")) %>% 
  filter(block != 0) %>% 
  filter(restk != "64") %>% # originally, we colledcted three levels of variaiblity, but the condition with kappa=64 showed similar results as kappa = 16
  as_data_frame() %>% 
  mutate(restk2 = recode(restk, `4` = "Brownian-like", `16` = "Ballistic-like"))

participants_exp1a <- read.csv(file.path(data_pth,"exp1a","participants_150526.csv"))

```

## Participant info
```{r participant info exp 1a}
table(participants_exp1a$gender)
c(mean(participants_exp1a$age), sd(participants_exp1a$age)) %>% round(2)
```

## Results

```{r exp1 results accuracy}
df_exp1a %>% 
  group_by(restk) %>% 
  summarize(accuracy_mean = mean(accuracy, na.rm = T)) %>% 
  knitr::kable(digits = 2)

df_exp1a %>% 
  group_by(restk,type) %>% 
  summarize(accuracy_mean = mean(accuracy, na.rm = T)) %>% 
  knitr::kable(digits = 2)

```

## Figure 1

```{r Figure 1}
traj_k64 <- read.table(file.path(data_pth, "exp1a", "trajectories", "T001_64_rest4.csv")) %>% select(x = V1,y = V2) %>% as_data_frame()
p1 <- traj_k64 %>% ggplot(aes(x,y)) + geom_point() + theme(aspect.ratio = 1) + xlim(-15,15)+ylim(-15,15) +
  ggtitle("Ballistic-like")

traj_k4 <- read.table(file.path(data_pth, "exp1a", "trajectories", "T001_4_rest4.csv")) %>% select(x = V3,y = V4) %>% as_data_frame()
p2 <- traj_k4 %>% ggplot(aes(x,y)) + geom_point() + theme(aspect.ratio = 1) + xlim(-15,15)+ylim(-15,15)+
  ggtitle("Brownian-like")
gridExtra::grid.arrange(p1,p2, ncol = 2)

ggsave(file.path(plot_pth, "Figure1_ballistic.png"), p1, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure1_brownian.png"), p2, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure1_ballistic.eps"), p1, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure1_brownian.eps"), p2, width = 6, height = 6)
```

## Figure 2

```{r Figure 2}

p1 <- plot_fig2(4, "Brownian-like-distractors")
p2 <- plot_fig2(16, "Ballistic-like-distractors")

gridExtra::grid.arrange(p1,p2, ncol = 2)

ggsave(file.path(plot_pth, "Figure2_brownian.png"), p1, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure2_ballistic.png"), p2, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure2_brownian.eps"), p1, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure2_ballistic.eps"), p2, width = 6, height = 6)
```

```{r exp1a - fitting weibull}

df.pf <- df_exp1a %>% 
  filter(restk2 == "Ballistic-like", type %in% c(2,4,8,16,32))

df.pf2 <- df.pf %>% 
  group_by(type) %>% 
  summarise(yes = sum(accuracy), no = n() - sum(accuracy), accuracy = 1 - mean(accuracy))

pf <- psyfun.2asym(cbind(yes, no) ~ type, data = df.pf2, link = weib.2asym, init.g = 0.125, init.lam = 0.05 )

x <- seq(-2,32,0.01)
y <- predict(pf, data.frame(type = x), type="response")

dfmodel <- data.frame(x, y)

p <- ggplot(data = dfmodel, aes(x = x, y = y)) + geom_line(color="red", size =1.5) + stat_summary(data = df.pf, aes(x = type, y=accuracy),fun.data = "mean_cl_boot", size = 1) + scale_x_continuous(expression("Target's"~kappa), breaks = df.pf2$type) + ylab("Accuracy")
p

```

## Overall differences

```{r exp 1a - overall differences, warning=F, message = F}
df_exp1a %>% 
  filter(restk2 == "Brownian-like") %>% 
  ez::ezANOVA(dv = accuracy,
        wid = subject,
        within = type, 
        detailed = T) %>% 
  apa::anova_apa()

df_exp1a %>% 
  filter(restk2 == "Ballistic-like") %>% 
  ez::ezANOVA(dv = accuracy,
        wid = subject,
        within = type, 
        detailed = T) %>% 
  apa::anova_apa()
```

## Tests with baseline

```{r exp 1a - test vs baseline}
df_exp1a %>% 
  group_by(type, restk,subject) %>% 
  summarize(accuracy = mean(accuracy)) %>% 
  group_by(restk, type) %>% 
  filter(accuracy < 1) %>% 
  do(t.test(.$accuracy, mu = .125) %>% 
       broom::tidy()) %>% 
  mutate(p.value = round(p.value, 3)) %>% 
  knitr::kable()
```

## Difference between symmetrical cases
```{r exp1a - difference between symmetrical case, warning=F, message = F}
df_exp1a %>% 
  filter((restk == 4 & type == 16) |(restk == 16 & type == 4)) %>% 
  ez::ezANOVA(
        dv = accuracy,
        wid = subject,
        within = restk, 
        detailed = T) %>% 
  apa::anova_apa()

```

# Experiment 1b

```{r exp1b  - load data}
df_exp1b <- readRDS(file.path(data_pth,"exp1b","responses","responses_150622.rds")) %>% 
  filter(block != 0)

participants_exp1b <- read.csv(file.path(data_pth,"exp1b","participants_150526.csv"), fileEncoding = "UTF8")

df2_exp1b <- rbind(df_exp1b,df_exp1b)
df2_exp1b$version <- 2
df2_exp1b$version[1:nrow(df_exp1b)] <- 1

# this is a little hack, just to make a correct ggplot with one line
tmp <- df_exp1b$restk
df2_exp1b$restk[df2_exp1b$version == 2] <- df_exp1b$type
df2_exp1b$type[df2_exp1b$version == 2] <- tmp

df2_exp1b$version <- factor(df2_exp1b$version, labels = c("fix distractors' k", "fix target's k"))
```

## Participant info
```{r}
table(participants_exp1b$gender)
c(mean(participants_exp1b$age), sd(participants_exp1b$age)) %>% round(2)
```

## Figure 3

```{r plots experiment 2}

p_k2 <- plot_fig3(2)
p_k4 <- plot_fig3(4)
p_k8 <- plot_fig3(8)
p_k16 <- plot_fig3(16)
p_k64 <- plot_fig3(64)
p_k64_legend <- plot_fig3(64, show_legend = T)

gridExtra::grid.arrange(p_k2,p_k4, p_k8, p_k16, p_k64, ncol = 3)
ggsave(file.path(plot_pth, "Fig3_k2.eps"), p_k2, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig3_k4.eps"), p_k4, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig3_k8.eps"), p_k8, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig3_k16.eps"), p_k16, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig3_k64.eps"), p_k64, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig3_k64_legend.eps"), p_k64_legend, width = 6, height = 6)
# final plot was created in inkscape
```

## Compute the asymmetry

We run multiple within subject aova

```{r compute the assymetry - exp1b, warning=F}
df3_exp1b <- df2_exp1b %>% group_by(type, restk) %>% do(p=compute.p(.))

df3_exp1b$p <- unlist(df3_exp1b$p) %>% round(3)
df4_exp1b <- df2_exp1b %>% group_by(type, restk) %>% do(ges=compute.des(.))
df4_exp1b$ges <- unlist(df4_exp1b$ges) %>% round(2)
df3_exp1b$ges <- df4_exp1b$ges

df5_exp1b <- df2_exp1b %>% group_by(type, restk, version) %>% summarize(accuracy = mean(accuracy)) %>% 
  group_by(type, restk) %>%
  summarize(acc_diff = accuracy[version == "fix distractors' k"] - accuracy[version == "fix target's k"])

m <- matrix(df3_exp1b$ges, nrow = 5)
rownames(m) <- levels(df_exp1b$type)
colnames(m) <- levels(df_exp1b$type)
m[upper.tri(m,diag = T)] <- NA

df3_exp1b$ges <- as.vector(m)

m <- matrix(df3_exp1b$p, nrow = 5)
m[upper.tri(m, diag = T)] <- NA
df3_exp1b$p <- as.vector(m)

m <- matrix(df5_exp1b$acc_diff, nrow = 5)
m[upper.tri(m, diag = T)] <- NA
df3_exp1b$acc_diff <- as.vector(m)

df3_exp1b$target_lower_than_dist <- df3_exp1b$acc_diff < 0
df3_exp1b %>% filter(target_lower_than_dist)

df3_exp1b %>% filter(p < .05)
```

## Test with baseline

```{r}
df_exp1b %>% 
  group_by(type, restk, subject) %>% 
  summarize(accuracy = mean(accuracy)) %>% group_by(restk, type) %>%
  filter(accuracy < 1) %>% 
  do(t.test(.$accuracy, mu = .125) %>% 
       broom::tidy()) %>% 
  mutate(p.value = round(p.value, 3)) %>% 
  knitr::kable()
```

# Experiment 2a

```{r load data - exp 2a}
df_exp2a <- readRDS(file.path(data_pth,"exp2a","responses","responses_150410.rds")) %>% 
  filter(block != 0) %>% 
  mutate(restk = restk * 100, type = type * 100) %>% 
  mutate(restk2 = recode(restk, `2` = "Slow-moving-distractors", `4` = "Fast-moving-distractors"))

participants_exp2a <- read.csv(file.path(data_pth,"exp2a","participants_171023.csv"))
```

## Participant info

```{r exp2a - participant info}
table(participants_exp2a$gender)
c(mean(participants_exp2a$age), sd(participants_exp2a$age)) %>% round(2)
```

## Results

```{r exp2a - results accuracy}
df_exp2a %>% 
  group_by(restk) %>% 
  summarize(accuracy_mean = mean(accuracy, na.rm = T)) %>% knitr::kable(digits = 2)
```

## Figure 4

```{r Figure 4}

p1 <- plot_fig4(2, ggtit = "Slow-moving-distractors")

p2 <- plot_fig4(4, ggtit = "Fast-moving-distractors")

gridExtra::grid.arrange(p1,p2, ncol = 2)

ggsave(file.path(plot_pth, "Figure4_slow.png"), p1, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure4_fast.png"), p2, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure4_slow.eps"), p1, width = 6, height = 6)
ggsave(file.path(plot_pth, "Figure4_fast.eps"), p2, width = 6, height = 6)
```

## Statistics

```{r exp 2a - statistics}
df_exp2a %>% 
  group_by(restk) %>% 
  summarize(acc = mean(accuracy) * 100, sd = sd(accuracy) * 100) %>% 
  knitr::kable()
```

## Difference between symmetrical cases
```{r exp 2a - symmetrical case, warning=F, message=F}

df_exp2a %>% 
  filter((restk == 2 & type == 4) |(restk == 4 & type == 2)) %>% 
  ez::ezANOVA(
        dv = accuracy,
        wid = subject,
        within = restk, 
        detailed = T) %>% 
  apa::anova_apa()

df_exp2a %>% 
  filter((restk == 2 & type == 4) |(restk == 4 & type == 2)) %>%
  ez::ezStats(
        dv = accuracy,
        wid = subject,
        within = restk)

```

## Test with baseline

```{r}
df_exp2a %>% group_by(type, restk,subject) %>% summarize(accuracy = mean(accuracy)) %>% group_by(restk, type) %>% filter(accuracy <1) %>% do(t.test(.$accuracy, mu = .125) %>% broom::tidy()) %>% mutate(p.value = round(p.value, 3)) %>% knitr::kable()
```

# Experiment 2b

```{r exp 2b - load data}
df_exp2b <- readRDS(here::here("..", "data", "exp2b", "responses", "responses_171023.rds")) %>% 
 filter(block != 0)

df_exp2b <- df_exp2b %>% mutate(restk = restk * 100, type = type * 100)

participants_exp2b <- read.csv(here::here("..", "data", "exp2b", "participants_171023.csv"))

df2_exp2b <- rbind(df_exp2b, df_exp2b)
df2_exp2b$version <- 2
df2_exp2b$version[1:nrow(df_exp2b)] <- 1

# this is a little hack, just to make a correct ggplot with one line
tmp <- df_exp2b$restk
df2_exp2b$restk[df2_exp2b$version == 2] <- df_exp2b$type
df2_exp2b$type[df2_exp2b$version == 2] <- tmp

df2_exp2b$version <- factor(df2_exp2b$version, labels = c("fix distractors' speed", "fix target's speed"))
```

## Participant info
```{r exp2b - participants info}
table(participants_exp2b$gender)
c(mean(participants_exp2b$age), sd(participants_exp2b$age)) %>% round(2)
```

## Figure 5

```{r Figure 5}

p_sp05 <- plot_fig5(0.5)
p_sp1 <- plot_fig5(1)
p_sp2 <- plot_fig5(2)
p_sp4 <- plot_fig5(4)
p_sp8 <- plot_fig5(8)
p_sp8_legend <- plot_fig5(8, show_legend = T)

gridExtra::grid.arrange(p_sp05,p_sp1, p_sp2, p_sp4, p_sp8, ncol = 3)
ggsave(file.path(plot_pth, "Fig5_sp05.eps"), p_sp05, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig5_sp1.eps"), p_sp1, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig5_sp2.eps"), p_sp2, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig5_sp4.eps"), p_sp4, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig5_sp8.eps"), p_sp8, width = 6, height = 6)
ggsave(file.path(plot_pth, "Fig5_sp8_legend.eps"), p_sp8_legend, width = 6, height = 6)
# final plot was created in inkscape
```

## Compute the asymmetry

```{r exp 2b - compute the assymetry, warning=F}
df3_exp2b <- df2_exp2b %>% group_by(type, restk) %>% do(p=compute.p(.))

df3_exp2b$p <- unlist(df3_exp2b$p) %>% round(3)
df4_exp2b <- df2_exp2b %>% group_by(type, restk) %>% do(ges=compute.des(.))
df4_exp2b$ges <- unlist(df4_exp2b$ges) %>% round(2)
df3_exp2b$ges <- df4_exp2b$ges

df5_exp2b <- df2_exp2b %>% group_by(type, restk, version) %>% summarize(accuracy = mean(accuracy)) %>% 
  group_by(type, restk) %>%
  summarize(acc_diff = accuracy[version == "fix distractors' speed"] - accuracy[version == "fix target's speed"])

m <- matrix(df3_exp2b$ges, nrow = 5)
rownames(m) <- levels(df_exp2b$type)
colnames(m) <- levels(df_exp2b$type)
m[upper.tri(m,diag = T)] <- NA

df3_exp2b$ges <- as.vector(m)

m <- matrix(df3_exp2b$p, nrow = 5)
m[upper.tri(m, diag = T)] <- NA
df3_exp2b$p <- as.vector(m)

m <- matrix(df5_exp2b$acc_diff, nrow = 5)
m[upper.tri(m, diag = T)] <- NA
df3_exp2b$acc_diff <- as.vector(m)

df3_exp2b$target_lower_than_dist <- df3_exp2b$acc_diff < 0
df3_exp2b %>% filter(target_lower_than_dist) %>% knitr::kable()

df3_exp2b %>% filter(p < 0.05) %>% knitr::kable()

differencies <- df3_exp2b %>% filter(target_lower_than_dist)  %>% .$acc_diff *-1
differencies %>% range() %>% round(2)

```