### 2018-09-25 R/manuscript_app_alt_clean.Rmd
changes according the paper changes

---
title: "manuscript - APP"
author: "Filip Dechterenko"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

## Load libraries

```{r load libraries, warning=F, message=F, echo=FALSE}

if (!require(tidyverse)) {install.packages("tidyverse"); library(tidyverse)}
if (!require(ez)) {install.packages("ez"); library(ez)}
if (!require(psyphy)) {install.packages("psyphy"); library(psyphy)}

if (!require(here)) {install.packages("here")}

library(lme4)
library(lmerTest)

theme_set(theme_bw(16))
source("tools.R")

data_pth <- here::here("..", "data")
```

```{r echo=FALSE}
# Exp 1b
df_exp1b <- 
  readRDS(
    file.path(data_pth, "exp1b", "responses", "responses_150622.rds")) %>% 
  filter(block != 0) %>% as_tibble(.) %>% 
  rename(targetk = type) %>% mutate(id = factor(subject)) %>% 
  mutate(contrast_abs = abs(log2(targetk / restk)), 
         target_straight = (targetk > restk),
         ratio = (besselI(restk,0) * exp(targetk)) / (besselI(targetk,0) * exp(restk)),
         ratio_abs = abs(log((besselI(restk,0) * exp(targetk))) / (besselI(targetk,0) * exp(restk))),
         ratio_abs = abs(log(
           (besselI(restk,0) * exp(targetk))) / (besselI(targetk,0) * exp(restk))))
participants_exp1b <- 
  read.csv(file.path(data_pth, "exp1b", "participants_150526.csv"), fileEncoding = "UTF8")

# Exp 2b
df_exp2b <- 
  readRDS(
    file.path(data_pth, "exp2b", "responses", "responses_171023.rds")) %>% 
 filter(block != 0) %>% as_tibble(.) %>% 
  rename(targetk = type) %>% mutate(id = factor(subject)) %>% 
  mutate(contrast_abs = abs(log2(targetk / restk)), 
         target_faster = (targetk > restk))
participants_exp2b <- 
  read.csv(file.path(data_pth, "exp2b", "participants_171023.csv"))
```

```{r kl distance}

ks <- c(2,4,8,16,64)
n <- length(ks)

m <- matrix(rep(0,n*n), ncol = n)
source("tools.R")
for (i in 1:length(ks)) {
  for (j in i:length(ks)) {
    m[i,j] <- kl_distance(ks[i],ks[j])    
  } 
}
rownames(m) <- colnames(m) <- ks

df_kl <- as_data_frame(m) %>%
  mutate(k1 = rownames(m)) %>% 
  gather(k2, value = kl_distance,-k1) %>% 
  mutate(k1 = as.numeric(k1), k2 = as.numeric(k2)) %>% 
  filter(k2>=k1) %>% 
  distinct()

df_kl <- df_kl %>% bind_rows(df_kl %>% mutate(tmp =k1, k1=k2,k2=tmp) %>% select(-tmp))

df_exp1b <- df_exp1b %>% 
  left_join(df_kl, by = c("targetk"="k1","restk"="k2"))
```

# Experiment 1b

```{r echo=F}
# accuracy summarised
accu_summary <- df_exp1b %>% 
  mutate(contrast_abs_f = as.character(contrast_abs)) %>% 
  group_by(id, contrast_abs, contrast_abs_f, target_straight, kl_distance,ratio,ratio_abs) %>% 
  dplyr::summarise(accu = mean(accuracy), n = n()) %>% 
  ungroup()
accu_summary_wo_0 <- accu_summary %>% filter(contrast_abs > 0)

# duplicated values for contrast 0
accu_summaryx <- bind_rows(
  accu_summary,
  accu_summary %>% 
    filter(contrast_abs == 0, target_straight == F) %>% 
    mutate(target_straight = T)
)

# raw data with duplicated values for contrast 0
df_exp1bx <- bind_rows(
  df_exp1b,
  df_exp1b %>% 
    filter(contrast_abs == 0, target_straight == F) %>% 
    mutate(target_straight = T)
)
# yes no for psychophysics
d1b <- df_exp1b %>% 
  group_by(id, target_straight, contrast_abs) %>% 
  summarise(yes = sum(accuracy), no = n() - yes, accuracy = mean(accuracy))
d1bx <- bind_rows(
  d1b,
  d1b %>% 
    filter(contrast_abs == 0) %>% ungroup() %>% mutate(target_straight = T)
)
```

## Individual data
```{r}
d1bx %>% ggplot(aes(x = contrast_abs, y = accuracy, colour = target_straight)) +
  geom_point(alpha = .5) + facet_wrap(~id)
```

```{r Charts1b}
#accu_summary %>% 
#  ggplot(aes(x = contrast_abs, y = accu, colour = target_straight)) + 
#  geom_jitter() 
accu_summaryx %>% 
  ggplot(
    aes(x = contrast_abs, y = accu, 
        group = target_straight, shape = target_straight, linetype = target_straight)) + 
  stat_summary(fun.data = "mean_cl_boot", alpha = .5) +
  stat_summary(fun.y = "mean", alpha = .5, geom = "line") + 
  theme(aspect.ratio = 1) + 
  scale_shape_discrete("Target straighter") + 
  xlab("contrast") + ylab("Accuracy")
```

```{r Charts1b kl distance}
#accu_summary %>% 
#  ggplot(aes(x = contrast_abs, y = accu, colour = target_straight)) + 
#  geom_jitter() 
accu_summaryx %>% 
  ggplot(
    aes(x = kl_distance, y = accu, 
        group = target_straight, shape = target_straight, linetype = target_straight)) + 
  stat_summary(fun.data = "mean_cl_boot", alpha = .5) +
  stat_summary(fun.y = "mean", alpha = .5, geom = "line") +

  theme(aspect.ratio = 1) + 
  xlab("kl distance") + ylab("Accuracy")
```

```{r Charts1b ratio}
#accu_summary %>% 
#  ggplot(aes(x = contrast_abs, y = accu, colour = target_straight)) + 
#  geom_jitter() 
accu_summaryx %>% 
  ggplot(
    aes(x = ratio, y = accu, 
        group = target_straight, shape = target_straight, linetype = target_straight)) + 
  stat_summary(fun.data = "mean_cl_boot", alpha = .5) +
  stat_summary(fun.y = "mean", alpha = .5, geom = "line") +

  theme(aspect.ratio = 1) + 
  xlab("PDF ratio") + ylab("Accuracy") +scale_x_log10()
```

```{r Charts1b ratio abs}
#accu_summary %>% 
#  ggplot(aes(x = contrast_abs, y = accu, colour = target_straight)) + 
#  geom_jitter() 
accu_summaryx %>% 
  ggplot(
    aes(x = ratio_abs, y = accu, 
        group = target_straight, shape = target_straight, linetype = target_straight)) + 
  stat_summary(fun.data = "mean_cl_boot", alpha = .5) +
  stat_summary(fun.y = "mean", alpha = .5, geom = "line") +

  theme(aspect.ratio = 1) + 
  xlab("ratio") + ylab("Accuracy") +scale_x_continuous(trans ="log")
```

```{r Charts1b without one participant}
accu_summary %>% 
  filter(id!= 13)
  ggplot(aes(x = contrast_abs, y = accu, colour = target_straight)) + 
  geom_jitter() 
accu_summary %>% 
  ggplot(
    aes(x = contrast_abs, y = accu, 
        group = target_straight, colour = target_straight)) + 
  stat_summary(fun.data = "mean_cl_boot", alpha = .5)
```

## ANOVA

### Linear 
```{r}
# linear
ezANOVA(accu_summaryx,
        accu, id, within = .(contrast_abs, target_straight))
```

### As factor
```{r}
# as factor and without contrast==0
ezANOVA(accu_summary_wo_0,
        accu, id, within = .(contrast_abs_f, target_straight),detailed = T)$ANOVA %>% 
  mutate(F = round(F,2), p = round(p,3),ges = round(ges,2),
         partial_eta = (SSn/(SSn+SSd)) %>% round(2))

```

### Post hoc tests

```{r}
(accu_summary_wo_0 %>% 
  filter(contrast_abs_f == 1) %>% 
ezANOVA(accu, id, within = .(target_straight),detailed = T))$ANOVA %>% 
  mutate(F = round(F,2), p = round(p,3),ges = round(ges,2),
         partial_eta = (SSn/(SSn+SSd)) %>% round(2))

(accu_summary_wo_0 %>% 
  filter(contrast_abs_f == 2) %>% 
ezANOVA(accu, id, within = .(target_straight),detailed = T))$ANOVA %>% 
  mutate(F = round(F,2), p = round(p,3),ges = round(ges,2),
         partial_eta = (SSn/(SSn+SSd)) %>% round(2))

(accu_summary_wo_0 %>% 
  filter(contrast_abs_f == 3) %>% 
ezANOVA(accu, id, within = .(target_straight),detailed = T))$ANOVA %>% 
  mutate(F = round(F,2), p = round(p,3),ges = round(ges,2),
         partial_eta = (SSn/(SSn+SSd)) %>% round(2))

(accu_summary_wo_0 %>% 
  filter(contrast_abs_f == 4) %>% 
ezANOVA(accu, id, within = .(target_straight),detailed = T))$ANOVA %>% 
  mutate(F = round(F,2), p = round(p,3),ges = round(ges,2),
         partial_eta = (SSn/(SSn+SSd)) %>% round(2))

```

## Mixed models

### Analysis as factors

```{r}
# as factor
m0 <- lmer(accu ~ (1|id), 
           data = accu_summary_wo_0)
summary(m0)
m1 <- lmer(accu ~ contrast_abs_f + (1|id), 
           data = accu_summary_wo_0)
summary(m1)
m2 <- lmer(accu ~ contrast_abs_f + target_straight + (1|id), 
           data = accu_summary_wo_0)
summary(m2)
m3 <- lmer(accu ~ contrast_abs_f * target_straight + (1|id), 
           data = accu_summary_wo_0)
summary(m3)
m4 <- lmer(accu ~ contrast_abs_f + contrast_abs_f:target_straight + (1|id), 
           data = accu_summary_wo_0)
```

### Linear

```{r}
m0 <- lmer(accu ~ (contrast_abs|id), data = accu_summaryx)
summary(m0)
m1 <- lmer(accu ~ contrast_abs + (contrast_abs|id), data = accu_summaryx)
summary(m1)
m2 <- lmer(accu ~ contrast_abs + target_straight + (contrast_abs|id), data = accu_summaryx)
summary(m2)
m3 <- lmer(accu ~ contrast_abs * target_straight + (1|id), data = accu_summaryx)
summary(m3)
m4 <- lmer(accu ~ contrast_abs + contrast_abs:target_straight + (1|id), data = accu_summaryx)
summary(m4)

```

### Logistic
```{r}
m0 <- glmer(accuracy ~ (1|id), 
            data = df_exp1bx, family = binomial())
summary(m0)
m1 <- glmer(accuracy ~ contrast_abs + (1|id), 
            data = df_exp1bx, family = binomial())
summary(m1)
m2 <- glmer(accuracy ~ contrast_abs + target_straight + (1|id), 
            data = df_exp1bx, family = binomial())
summary(m2)
m3 <- glmer(accuracy ~ contrast_abs * target_straight + (1|id), 
            data = df_exp1bx, family = binomial())
summary(m3)

```

### Psychophysics

```{r}
df.pf <- df_exp1bx %>% 
  group_by(contrast_abs, target_straight) %>% 
  summarise(yes = sum(accuracy), no = n() - sum(accuracy), 
            accuracy = 1 - mean(accuracy))

pf_ts <- psyfun.2asym(cbind(yes, no) ~ contrast_abs, 
                      data = df.pf %>% filter(target_straight), 
                      link = weib.2asym, init.g = 0.125, init.lam = 0.05 )
pf_tb <- psyfun.2asym(cbind(yes, no) ~ contrast_abs, 
                      data = df.pf %>% filter(!target_straight), 
                      link = weib.2asym, init.g = 0.125, init.lam = 0.05 )

x <- seq(0, 6,0.01)
ys <- predict(pf_ts, data.frame(contrast_abs = x), type = "response")
yb <- predict(pf_tb, data.frame(contrast_abs = x), type = "response")

dfmodel <- data.frame(x, y = c(ys, yb), 
                      target_straight = rep(c(T,F), each = length(x)))

p <- ggplot(data = dfmodel, aes(x = x, y = y, colour = target_straight)) + 
  geom_line(size = 1.5) + 
  stat_summary(data = df.pf, aes(x = contrast_abs, y = accuracy), 
               fun.data = "mean_cl_boot", size = 1) +
  NULL
p

summary(pf_ts)
summary(pf_tb)
```

# Experiment 2b

```{r echo=F}
# accuracy summarised
accu_summary <- df_exp2b %>% 
  mutate(contrast_abs_f = as.character(contrast_abs)) %>% 
  group_by(id, contrast_abs, contrast_abs_f, target_faster) %>% 
  dplyr::summarise(accu = mean(accuracy), n = n()) 
accu_summary_wo_0 <- accu_summary %>% filter(contrast_abs > 0)

# duplicated values for contrast 0
accu_summaryx <- bind_rows(
  accu_summary,
  accu_summary %>% 
    filter(contrast_abs == 0, target_faster == F) %>% 
    mutate(target_faster = T)
)

# raw data with duplicated values for contrast 0
df_exp2bx <- bind_rows(
  df_exp2b,
  df_exp2b %>% 
    filter(contrast_abs == 0, target_faster == F) %>% 
    mutate(target_faster = T)
)
# yes no for psychophysics
d2b <- df_exp2b %>% 
  group_by(id, target_faster, contrast_abs) %>% 
  summarise(yes = sum(accuracy), no = n() - yes, accuracy = mean(accuracy))
d2bx <- bind_rows(
  d2b,
  d2b %>% 
    filter(contrast_abs == 0) %>% ungroup() %>% mutate(target_faster = T)
)
```

## Individual data
```{r}
d2bx %>% ggplot(aes(x = contrast_abs, y = accuracy, colour = target_faster)) +
  geom_point(alpha = .5) + facet_wrap(~id)
weird_performance <- c(1, 5, 14)
```

```{r Charts2b}
accu_summary %>% 
  ggplot(aes(x = contrast_abs, y = accu, colour = target_faster)) + 
  geom_jitter() 
accu_summary %>% 
  ggplot(
    aes(x = contrast_abs, y = accu, 
        group = target_faster, colour = target_faster)) + 
  stat_summary(fun.data = "mean_cl_boot", alpha = .5)
```

```{r Charts2b without weird}
accu_summary %>% 
  filter(!id %in% weird_performance) %>% 
  ggplot(
    aes(x = contrast_abs, y = accu, 
        group = target_faster, colour = target_faster)) + 
  stat_summary(fun.data = "mean_cl_boot", alpha = .5)
```

## ANOVA

### Linear 
```{r}
# linear
ezANOVA(accu_summaryx,
        accu, id, within = .(contrast_abs, target_faster))
```

### As factor
```{r}
# as factor and without contrast==0
ezANOVA(accu_summary_wo_0,
        accu, id, within = .(contrast_abs_f, target_faster))
```

### Test without weird performance
```{r}
# linear
ezANOVA(accu_summaryx %>% filter(!id %in% weird_performance),
        accu, id, within = .(contrast_abs, target_faster))
# as factor and without contrast==0
ezANOVA(accu_summary_wo_0 %>% filter(!id %in% weird_performance),
        accu, id, within = .(contrast_abs_f, target_faster))
```

## Mixed models

### Analysis as factors

```{r}
# as factor
m0 <- lmer(accu ~ (1|id), 
           data = accu_summary_wo_0)
summary(m0)
m1 <- lmer(accu ~ contrast_abs_f + (1|id), 
           data = accu_summary_wo_0)
summary(m1)
m2 <- lmer(accu ~ contrast_abs_f + target_faster + (1|id), 
           data = accu_summary_wo_0)
summary(m2)
m3 <- lmer(accu ~ contrast_abs_f * target_faster + (1|id), 
           data = accu_summary_wo_0)
summary(m3)
m4 <- lmer(accu ~ contrast_abs_f + contrast_abs_f:target_faster + (1|id), 
           data = accu_summary_wo_0)
```

### Linear

```{r}
m0 <- lmer(accu ~ (contrast_abs|id), data = accu_summaryx)
summary(m0)
m1 <- lmer(accu ~ contrast_abs + (contrast_abs|id), data = accu_summaryx)
summary(m1)
m2 <- lmer(accu ~ contrast_abs + target_faster + (contrast_abs|id), data = accu_summaryx)
summary(m2)
m3 <- lmer(accu ~ contrast_abs * target_faster + (1|id), data = accu_summaryx)
summary(m3)
m4 <- lmer(accu ~ contrast_abs + contrast_abs:target_faster + (1|id), data = accu_summaryx)
summary(m4)

```

### Logistic
```{r}
m0 <- glmer(accuracy ~ (1|id), 
            data = df_exp2bx, family = binomial())
summary(m0)
m1 <- glmer(accuracy ~ contrast_abs + (1|id), 
            data = df_exp2bx, family = binomial())
summary(m1)
m2 <- glmer(accuracy ~ contrast_abs + target_faster + (1|id), 
            data = df_exp2bx, family = binomial())
summary(m2)
m3 <- glmer(accuracy ~ contrast_abs * target_faster + (1|id), 
            data = df_exp2bx, family = binomial())
summary(m3)
```

### Test w/o weird performance in logistic regression
```{r}
m0 <- glmer(accuracy ~ (1|id), 
            data = df_exp2bx %>% filter(!id %in% weird_performance), family = binomial())
summary(m0)
m3 <- glmer(accuracy ~ contrast_abs * target_faster + (1|id), 
            data = df_exp2bx %>% filter(!id %in% weird_performance), family = binomial())
summary(m3)
anova(m3,m0)
```

### Psychophysics

```{r}
df.pf <- df_exp2bx %>% 
  group_by(contrast_abs, target_faster) %>% 
  summarise(yes = sum(accuracy), no = n() - sum(accuracy), 
            accuracy = 1 - mean(accuracy))

pf_ts <- psyfun.2asym(cbind(yes, no) ~ contrast_abs, 
                      data = df.pf %>% filter(target_faster), 
                      link = weib.2asym, init.g = 0.125, init.lam = 0.05)
pf_tb <- psyfun.2asym(cbind(yes, no) ~ contrast_abs, 
                      data = df.pf %>% filter(!target_faster), 
                      link = weib.2asym, init.g = 0.125, init.lam = 0.05)

x <- seq(0, 6,0.01)
ys <- predict(pf_ts, data.frame(contrast_abs = x), type = "response")
yb <- predict(pf_tb, data.frame(contrast_abs = x), type = "response")

dfmodel <- data.frame(x, y = c(ys, yb), 
                      target_faster = rep(c(T,F), each = length(x)))

p <- ggplot(data = dfmodel, aes(x = x, y = y, colour = target_faster)) + 
  geom_line(size = 1.5) + 
  stat_summary(data = df.pf, aes(x = contrast_abs, y = accuracy), 
               fun.data = "mean_cl_boot", size = 1) +
  NULL
p

summary(pf_ts)
summary(pf_tb)
```